---
title: "CDS ODS reliability"
date: "August 2, 2021"
output:
  html_document: 
   toc: true
   toc_float: true
---

```{r setup, include=FALSE}

library(tidyverse)
library(readxl)
library(irr) # kappa
library(psych) # kappa

```


# Data prep

<!-- ## read in data -->
<!-- ```{r} -->
<!-- # wf -->
<!-- wf <- read_excel("data_WF_withNAPS_ir_reliability.xlsx", sheet = "data_WF_withNAPS_ir_reliability") %>% -->
<!--   dplyr::select(c(interrater, id, months_age, time, dur_min, AWC, CTC, CVC, cds_ohs)) %>% -->
<!--   filter(interrater == 1) %>% -->
<!--   mutate(study = "WF", -->
<!--          language = "spanish") %>% -->
<!--   rename("cds_ods_coder1" = "cds_ohs", -->
<!--          "dur_5min" = "dur_min", "AWC_5min" = "AWC", "CTC_5min" = "CTC", "CVC_5min" = "CVC") -->

<!-- # sot - outreach -->
<!-- sot_outreach <- read_excel("data_SOT_Outreach_withNAPS_ir_reliability.xlsx", sheet = "data_SOT_Outreach_withNAPS") %>% -->
<!--   dplyr::select(c(interrater, id, months_age, time, dur_min, AWC, CTC, CVC, cds_ohs)) %>% -->
<!--   filter(interrater == 1) %>% -->
<!--   mutate(study = "SOT_Outreach", -->
<!--          language = "english") %>% -->
<!--   rename("cds_ods_coder1" = "cds_ohs", -->
<!--          "dur_5min" = "dur_min", "AWC_5min" = "AWC", "CTC_5min" = "CTC", "CVC_5min" = "CVC") -->

<!-- # sot - stanford -->
<!-- sot_stanford <- read_excel("data_SOT_Stanford_withNAPS_ir_reliability.xlsx", sheet = "data_SOT_Stanford_withNAPS") %>% -->
<!--   dplyr::select(c(interrater, id, months_age, time, dur_min, AWC, CTC, CVC, cds_ohs)) %>% -->
<!--   filter(interrater == 1) %>% -->
<!--   mutate(study = "SOT_Stanford", -->
<!--          language = "english") %>% -->
<!--   rename("cds_ods_coder1" = "cds_ohs", -->
<!--          "dur_5min" = "dur_min", "AWC_5min" = "AWC", "CTC_5min" = "CTC", "CVC_5min" = "CVC") -->


<!-- # contx -->
<!-- contx <- read_excel("data_contx_5min_and_10min_withNAPS_200820_ir_reliability.xlsx", sheet = "data_contx_5min_and_10min_withN") %>% -->
<!--   dplyr::select(c(interrater, id, months_age, time, dur_5min, AWC_5min, CTC_5min, CVC_5min, cds_ohs, language)) %>% -->
<!--   filter(interrater == 1) %>% -->
<!--   mutate(study = "CONTX") %>% -->
<!--   rename("cds_ods_coder1" = "cds_ohs") -->

<!-- # combine all studies -->
<!-- first_coders <- rbind(wf, sot_outreach, sot_stanford, contx) %>% -->
<!--   dplyr::select(-interrater) -->
<!-- write.csv(first_coders, "./data_reliability/first_coders.csv") -->
<!-- ``` -->



# Combine first and second coder info
## create column with second coder criteria same for all studies
```{r}
# read in first coder data
first_coders <- read.csv("./data_reliability/first_coders.csv") %>% 
  dplyr::select(-X)

# create coder 2 cds_ods column and combine with first coder
coders <- read_excel("./data_reliability/CDS_ODS_tracking_revised_Sep52021.xlsx", sheet = "assigned") %>% 
  rename("id" = "ID", "dur_5min" = "duration", "AWC_5min" = "AWC", "LENA_on_child" = "LENA on child") %>% 
  mutate(cds_ods_coder2 = 
                   ifelse(CDS_pct_of_AWC == "true 0", "0",
                   ifelse(CDS_pct_of_AWC == "< 50%", "0", 
                   ifelse(CDS_pct_of_AWC == "50-70%", 1, 
                   ifelse(CDS_pct_of_AWC == "70% +", 1, "CHECK"))))) %>% 
  full_join(first_coders, by = c("id", "study", "language", "time", "AWC_5min", "dur_5min"))
str(coders)


# number of rows matches each dataset
coders_nrow <- coders %>% 
  group_by(study) %>% 
  count()
coders_nrow
```



## create column for different second coder cds_ods coding for contx vs. wf + sots
```{r}
# contx only
contx_firstcoder <- first_coders %>% 
  filter(study == "CONTX")


# create coder 2 cds_ods column and combine with first coder
contx <- read_excel("./data_reliability/CDS_ODS_tracking_revised_Sep52021.xlsx", sheet = "assigned") %>% 
  rename("id" = "ID", "dur_5min" = "duration", "AWC_5min" = "AWC", "LENA_on_child" = "LENA on child") %>% 
  filter(study == "CONTX") %>% 
  mutate(cds_ods_coder2 = 
                   ifelse(CDS_pct_of_AWC == "true 0", "0",
                   ifelse(CDS_pct_of_AWC == "< 50%", "0", 
                   ifelse(CDS_pct_of_AWC == "50-70%", 0, 
                   ifelse(CDS_pct_of_AWC == "70% +", 1, "CHECK"))))) %>% 
  full_join(contx_firstcoder, by = c("id", "study", "language", "time", "AWC_5min", "dur_5min"))


# wf and sot matching info
first_coders_wf_sots <- first_coders %>% 
  filter(study != "CONTX")


# create df with contx, wf and sots
coders_v2 <- read_excel("./data_reliability/CDS_ODS_tracking_revised_Sep52021.xlsx", sheet = "assigned") %>% 
  filter(study != "CONTX") %>% 
  rename("id" = "ID", "dur_5min" = "duration", "AWC_5min" = "AWC", "LENA_on_child" = "LENA on child") %>% 
  mutate(cds_ods_coder2 = 
                   ifelse(CDS_pct_of_AWC == "true 0", "0",
                   ifelse(CDS_pct_of_AWC == "< 50%", "0", 
                   ifelse(CDS_pct_of_AWC == "50-70%", 1, 
                   ifelse(CDS_pct_of_AWC == "70% +", 1, "CHECK"))))) %>% 
  full_join(first_coders_wf_sots, by = c("id", "study", "language", "time", "AWC_5min", "dur_5min")) %>% 
  rbind(contx) %>% 
  rename("cds_ods_coder2_v2" = "cds_ods_coder2") %>% 
  dplyr::select(c(study, language, id, rectime, time, cds_ods_coder2_v2))


```


## create final df with two versions of cds_ods coding
```{r}
coders <- coders %>% 
  full_join(coders_v2, by = c("id", "study", "language", "rectime", "time")) %>% 
  dplyr::select(c(id, study, language, rectime, time, dur_5min, AWC_5min, CDS_pct_of_AWC, LENA_on_child, 
                  cds_ods_coder1, cds_ods_coder2, cds_ods_coder2_v2))

```


## exclude data: 1) splits [from rater1 coding] and 2) LENA off child for > 50%
```{r}
# exclude data
coders_final <- coders %>% 
  filter(cds_ods_coder1 != "split") %>% 
  filter(LENA_on_child != "no >50%")
  
# view excludes
# splits
df_split <- coders %>% 
  filter(cds_ods_coder1 == "split")

nrow(df_split)
df_split %>% group_by(study) %>% count()

# pct excluded
pct_split <- (nrow(df_split)/nrow(coders))*100
pct_split

# lena less than 50%
df_lena_off_more_50pct <- coders %>% 
  filter(LENA_on_child == "no >50%")

nrow(df_lena_off_more_50pct)
df_lena_off_more_50pct %>% group_by(study) %>% count()

# pct excluded
pct_lenaoff50 <- (nrow(df_lena_off_more_50pct)/nrow(coders))*100
pct_lenaoff50

# total excluded
pct_split + pct_lenaoff50


# table
table(coders$cds_ods_coder1, coders$CDS_pct_of_AWC) 
# 'split' segments are indeed mostly 50%+ CDS (25 / 37; 68%)
table(coders$cds_ods_coder1, coders$CDS_pct_of_AWC)/ rowSums(table(coders$cds_ods_coder1, coders$CDS_pct_of_AWC))

```



# Reliability

## final data details
```{r}
# total kept
nrow(coders_final)
100 - (pct_split + pct_lenaoff50)

# count by study
coders_final_nrow <- coders_final %>% 
  group_by(study) %>% 
  count() %>% 
  full_join(coders_nrow, by = "study") %>% 
  rename("nrow_final" = "n.x", "nrow_full" = "n.y") %>% 
  mutate(prop_final = nrow_final/nrow_full)
coders_final_nrow

```


## coders_v1: proportion of matched codes
```{r}
# create match column
coders_final <- coders_final %>% 
  mutate(cds_ods_coder1 = as.numeric(cds_ods_coder1), 
         cds_ods_coder2 = as.numeric(cds_ods_coder2), 
         cds_ods_coder2_v2 = as.numeric(cds_ods_coder2_v2)) %>% 
  mutate(match_v1 = ifelse(cds_ods_coder1 == cds_ods_coder2, "match", "mismatch"), 
         match_v2 = ifelse(cds_ods_coder1 == cds_ods_coder2_v2, "match", "mismatch"))

# count of match for version 1
coders_final %>% 
  group_by(match_v1) %>% 
  count() %>% 
  mutate(percent = n/nrow(coders_final))


# count of match for version 2
coders_final %>% 
  group_by(match_v2) %>% 
  count() %>% 
  mutate(percent = n/nrow(coders_final))


# count of matches vs. mismatch - by STUDY
prop_v1 <- coders_final %>% 
  group_by(study, match_v1) %>% 
  count() %>% 
  group_by(study) %>% 
  mutate(study_n = sum(n)) %>% 
  mutate(percent = n/study_n) %>% 
  mutate(study = factor(study, levels = c("WF", "SOT_Outreach", "SOT_Stanford", "CONTX")))
prop_v1

prop_v2 <- coders_final %>% 
  group_by(study, match_v2) %>% 
  count() %>% 
  group_by(study) %>% 
  mutate(study_n = sum(n)) %>% 
  mutate(percent = n/study_n) %>% 
  mutate(study = factor(study, levels = c("WF", "SOT_Outreach", "SOT_Stanford", "CONTX")))
prop_v2

ggplot(prop_v1, aes(study, percent, fill = match_v1)) +
  geom_col(position = "dodge") + theme_classic()

ggplot(prop_v2, aes(study, percent, fill = match_v2)) +
  geom_col(position = "dodge") + theme_classic()


ggplot(coders_final, aes(AWC_5min, fill = study)) + 
  geom_histogram() +
  facet_wrap(~ match_v2, ncol = 1) + theme_classic()

```



## dfs of mismatch only in v2
```{r}
# CONTX
contx_mismatch_v1 <- coders_final %>% 
  filter(study == "CONTX" & match_v1 == "mismatch")

contx_mismatch_v2 <- coders_final %>% 
  filter(study == "CONTX" & match_v2 == "mismatch")

# SOT
sot_outreach_mismatch_v2 <- coders_final %>% 
  filter(study == "SOT_Outreach" & match_v2 == "mismatch")

sot_stanford_mismatch_v2 <- coders_final %>% 
  filter(study == "SOT_Stanford" & match_v2 == "mismatch")

# WF
wf_mismatch_v2 <- coders_final %>% 
  filter(study == "WF" & match_v2 == "mismatch")

contx_mismatch_v2
sot_outreach_mismatch_v2
sot_stanford_mismatch_v2
wf_mismatch_v2

```



## ALL data
```{r}
coders_final_kappa_v1 <- coders_final %>% 
  dplyr::select(c(cds_ods_coder1, cds_ods_coder2))

coders_final_kappa_v2 <- coders_final %>% 
  dplyr::select(c(cds_ods_coder1, cds_ods_coder2_v2))

kappa2(coders_final_kappa_v1)
kappa2(coders_final_kappa_v2)
```


## no CONTX
```{r}
coders_final_kappa_noCONTX <- coders_final %>% 
  filter(study != "CONTX") %>% 
  dplyr::select(c(cds_ods_coder1, cds_ods_coder2))
  
kappa2(coders_final_kappa_noCONTX)
```


# Reliability - per dataset
## WF only
```{r}
# we do not want weighted kappa; this takes into account disagreements, but weighted kappa is more relevant when data are rank ordered
coders_final_kappa_wf <- coders_final %>% 
  filter(study == "WF") %>% 
  dplyr::select(c(cds_ods_coder1, cds_ods_coder2))
  
kappa2(coders_final_kappa_wf)
```


## SOT Outreach only
```{r}
coders_final_kappa_sotOutreach <- coders_final %>% 
  filter(study == "SOT_Outreach") %>% 
  dplyr::select(c(cds_ods_coder1, cds_ods_coder2))
  
kappa2(coders_final_kappa_sotOutreach)
```


## SOT Stanford only
```{r}
coders_final_kappa_sotStanford <- coders_final %>% 
  filter(study == "SOT_Stanford") %>% 
  dplyr::select(c(cds_ods_coder1, cds_ods_coder2))
  
kappa2(coders_final_kappa_sotStanford)
```


## CONTX only
```{r}
coders_final_kappa_CONTX_v1 <- coders_final %>% 
  filter(study == "CONTX") %>% 
  dplyr::select(c(cds_ods_coder1, cds_ods_coder2))

coders_final_kappa_CONTX_v2 <- coders_final %>% 
  filter(study == "CONTX") %>% 
  dplyr::select(c(cds_ods_coder1, cds_ods_coder2_v2))
  
kappa2(coders_final_kappa_CONTX_v1)
kappa2(coders_final_kappa_CONTX_v2)
```


## Compare to Model Predictions

```{r}
# XGboost predictions from raw LENA features
#xgb_pred <- read_csv("rawLENAxgb.csv") # all 12931
xgb_pred <- read_csv("paper/xgb_preds_nonaps.csv")
# random forest predictions
#rf_pred <- read_csv("rf-predicted.csv") # only 9808 obs (the train data?)


# join by subject id, language, time (match study and age, too?)
hum_vs_mod <- coders_final %>% left_join(xgb_pred)

# model's binary predictions vs. original coders
kappa2(hum_vs_mod %>% dplyr::select(cds_ods_coder1, CDSpred)) # .5  gold standard ?

# model's binary predictions vs. new reliability coding
kappa2(hum_vs_mod %>% dplyr::select(cds_ods_coder2, CDSpred)) # .38

# reliability of all 3 (2 human, 1 model):
kappam.fleiss(hum_vs_mod %>% dplyr::select(cds_ods_coder1, cds_ods_coder2, CDSpred)) # .47

# reliability on just the 137/378 segments that the model has high confidence on:
high_conf <- hum_vs_mod %>% filter(xgb_pred < .25 | xgb_pred > .75)
kappam.fleiss(high_conf %>% dplyr::select(cds_ods_coder1, cds_ods_coder2, CDSpred)) # .68


# ToDo: ordinally-ranked %CDS in humans vs. model
hum_vs_mod$CDS_pct_of_AWC <- factor(hum_vs_mod$CDS_pct_of_AWC, 
                                       levels=c("true 0","< 50%" , "50-70%" ,"70% +"))
hum_vs_mod$CDS_pct_of_AWC <- as.numeric(hum_vs_mod$CDS_pct_of_AWC)
with(hum_vs_mod, cor.test(CDS_pct_of_AWC, xgb_pred, method="spearman")) # .49
#with(hum_vs_mod, plot(jitter(CDS_pct_of_AWC), xgb_pred)) 

# reliability of model with original coders across entire dataset
kappa2(xgb_pred %>% dplyr::select(CDSpred,cds_ohs)) # .42 
```

